---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(rvoronoi)
```

# rvoronoi

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
![](https://img.shields.io/badge/status-dangerously_unstable-red.svg)
[![R-CMD-check](https://github.com/coolbutuseless/rvoronoi/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/rvoronoi/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`rvoronoi` is a testing ground for some rendering ideas using fast delaunay triangulation
and voronoi tesselation.

For small sets of seed points (e.g. 20 points) this package can be 10x faster than 
`RTriangle` package.  For larger sets of seed points, the gap closes up - 
e.g. for N = 1000, still 2x faster than `RTriangle` package.

The core of this package is Steven Fortune's original C source code for his
sweep algorithm.   This code has been updated and adapted to run within R.
[Original source code (packaged as a shell archive)](https://netlib.sandia.gov/voronoi/sweep2)


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/rvoronoi) with:

``` r
# install.packages('remotes')
remotes::install_github('coolbutuseless/rvoronoi')
```


## Voronoi Tesselation

The following code calculates the voronoi tesselation on 20 random points.

Only finite length segments are plotted.  Segments which head to infinity 
are not plotted here.

```{r voronoi}
library(rvoronoi)

set.seed(2)
x <- runif(20)
y <- runif(20)

vor <- voronoi(x, y) 

# Plot the seed points
plot(x, y, asp = 1, col = 'red')

# Plot all finite segments.  
# This will not plot any of the segments which do not converge

fseg <- subset(vor$segment, v1 > 0 & v2 > 0)

segments(
  vor$vertex$x[ fseg$v1 ],
  vor$vertex$y[ fseg$v1 ],
  vor$vertex$x[ fseg$v2 ],
  vor$vertex$y[ fseg$v2 ],
)
```


## Delaunay Triangulation

```{r delaunay}
library(rvoronoi)

set.seed(2024)
x <- runif(10)
y <- runif(10)

del <- delaunay(x, y)$segment

# Plot the seed points
plot(x, y, asp = 1, col = 'red')

# Plot all finite segments.  
# This will not plot any of the segments which do not converge
segments(x[del$v1], y[del$v1], x[del$v2], y[del$v2])
segments(x[del$v3], y[del$v3], x[del$v2], y[del$v2])
segments(x[del$v1], y[del$v1], x[del$v3], y[del$v3])
```


## Delaunay Benchmark

Simple benchmark comparing this package with `{RTriangle}`

Benchmarking code for other packages is welcomed!

```{r benchmark}
library(RTriangle)

set.seed(2024)
N <- 20
x <- runif(N)
y <- runif(N)

del_rtriangle <- RTriangle::triangulate(RTriangle::pslg(P = cbind(x, y)))$T
del_new       <- delaunay(x, y)$segment

identical(
  rvoronoi:::normalise_del(del_new),
  rvoronoi:::normalise_del(del_rtriangle)
)
```


```{r echo=FALSE}
res <- bench::mark(
  del_rtriangle = triangulate(pslg(P = cbind(x, y)))$T,
  del_new       = delaunay(x, y),
  check = FALSE,
) 

res[1:5] |> knitr::kable()
```



## Debug plotting

* Used for debugging
* Not ready for general use

```{r}
set.seed(3)
N <- 10
x0 <- runif(N)
y0 <- runif(N)

vor <- voronoi(x0, y0)

if (FALSE) {
  ps <- extract_polygons_r(vor)
  ps
}

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red')
```


## Pathological 1

* 100 points in a circle

```{r}
theta <- seq(0, 2*pi, length.out = 101)[-1]
x <- cos(theta)
y <- sin(theta)

vor <- voronoi(x, y)

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red')
points(x, y)
```




## Pathological 2

* 100 points in a circle
* 1 point at the centre

```{r}
theta <- seq(0, 2*pi, length.out = 101)[-1]
x <- c(0, cos(theta))
y <- c(0, sin(theta))

vor <- voronoi(x, y)

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red')
points(x, y, pch = '+')
```

## Pathological 2a

* 2 concentric circles
* 1 point at the centre

```{r}
theta <- seq(0, 2*pi, length.out = 100)[-1]
x <- c(0, cos(theta), 2 * cos(theta))
y <- c(0, sin(theta), 2 * sin(theta))

vor <- voronoi(x, y)

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red')
points(x, y, pch = '+')
```


## Pathological 3

* 100 points in a line

```{r}
x <- seq(0, 2*pi, length.out = 100)
y <- 0.5 * x  

vor <- voronoi(x, y)

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red')
points(x, y, pch = '+')
```



## Pathological 4

* 2 concentric cirles (50 points each)
* 1 point at the centre

```{r}
theta <- seq(0, 2*pi, length.out = 4)[-1]
x <- c(0, cos(theta), 2 * cos(theta))
y <- c(0, sin(theta), 2 * sin(theta))

vor <- voronoi(x, y)

plot_vor(vor) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red') |>
  draw_vertices()
points(x, y, pch = '+')

if (FALSE) {
  plot(x, y, ann = F, axes = F, asp = 1)
  
  cols <- rainbow(length(vor$polygons))
  
  for (i in seq_along(vor$polygons)) {
  polygon(vor$polygons[[i]], col = cols[i])  
  text(x[i], y[i], labels = i)
  # points(x[i], y[i], col = 'black', pch = '+')
  }
  
}


```


```{r}
library(RTriangle)
theta <- seq(0, 2*pi, length.out = 6)[-1]
x <- c(0, cos(theta), 2 * cos(theta))
y <- c(0, sin(theta), 2 * sin(theta))

tri <- triangulate(pslg(P = cbind(x, y)))

v1 <- tri$VE[,1]
v2 <- tri$VE[,2]

line <- rep(1L, length(v1))

seg <- merge_vertices(tri$VP[,1], tri$VP[,2], line, v1, v2)

polys <- extract_polygons(tri$VP[,1], tri$VP[,2], seg$v1, seg$v2)

cols <- rainbow(length(polys))
plot(tri$VP, asp = 1, ann = F, axes = FALSE)
for (i in seq_along(polys)) {
  polygon(polys[[i]], col = cols[i])
}

```


## Random

```{r}
set.seed(1)
N <- 6
x <- runif(N)
y <- runif(N)

vor <- voronoi(x, y)

plot_vor(vor, buffer = 1) |>
  draw_segments() |> 
  draw_inf_lines() |>
  draw_inf_segments(col = 'hotpink') |>
  draw_bounded_polygons(border = 'red') |>
  draw_vertices()
points(x, y, pch = '+')
```


# Algorithms

* Fortune's Sweep Algorithm for Delaunay
    * Adapted original source code to call from R
* Polygon reconstruction
    * Use voronoi vertices and edge connectivity to reconstruct polygonal voronoi tiles
    * Using "An optimal algorithm for extracting the regions of a plane graph" by
      Jiang & Bunke, Pattern Recognition Letters 14 (1993) pp553-558.
    * Adapted algorithm from binary search to instead use pre-indexed search bounds.
* Point in polygon (to match seed points to voronoi polygons)
    * Search for bounding box collision to filter polygon candidates for 
      exhaustive testing
    * Optimised point-in-polygon search with early termination possible (as voronoi 
      polygons are **always** convex)
      Test edges using the "leftOf()" operator, and exit as soon as any 
      vertices are right of a polygon edge.
    * No benefit in building an acceleration structure for point-in-bounding-box
      when there of the order of ~100-1000 bounding boxes.  A linear search 
      (with flagging of polygons already claimed) is fast enough.
      If use case if voronoi of >> 1000 points, then time taken to build acceleration structure
      (e.g. hierachical bounding boxes) would become worth it.

## Related 

* [mdsumner's helpful gist](https://gist.github.com/mdsumner/8db5ac01e47fa86f10e7ebc372e0ebda?permalink_comment_id=5171205)
* [`RTriangle::triangulate()`]()
* [`terra::voronoi()`](https://rspatial.github.io/terra/reference/voronoi.html)
* [`geos::geos_voronoi_polygons()`]()

