---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width =  6,
  fig.height = 4
)

library(rvoronoi)
```

# rvoronoi

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
![](https://img.shields.io/badge/API-stabilising-lightblue.svg)
[![R-CMD-check](https://github.com/coolbutuseless/rvoronoi/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/rvoronoi/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`rvoronoi` provides fast Delaunay triangulation and Voronoi tessellation using
[Fortune's Sweep Line Algorithm](https://en.wikipedia.org/wiki/Fortune%27s_algorithm).

A key benefit of this package (besides *speed!*) is that the Voronoi tesselation 
can return a full set of enclosed polygons matched to the input sites. This 
makes it convenient for plotting.

The core of this package is Steven Fortune's original C source code for his
sweep algorithm.   This code has been updated and adapted to run within R.
[Original source code (packaged as a shell archive)](https://netlib.sandia.gov/voronoi/sweep2)

For small sets of site (e.g. 20 points) this package can be 10x faster than 
`RTriangle` package.  For larger sets of sites, the gap closes up - 
e.g. for N = 1000, still 2x faster than `RTriangle` package.


## Installation

You can install from [GitHub](https://github.com/coolbutuseless/rvoronoi) with:

``` r
# install.packages('remotes')
remotes::install_github('coolbutuseless/rvoronoi')
```

# In fair Voronoi, where we lay our scene

A Voronoi tessellation of a scene from 1996 movie [Romeo + Juliet](https://www.imdb.com/title/tt0117509/?ref_=fn_al_tt_1)

```{r romeo, fig.height = 4}
library(nara)
library(grid)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Load image
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
rj <- jpeg::readJPEG("man/figures/rj.jpg", native = TRUE)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Sample lots of random points in the image
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
set.seed(5)
N <- 4000
x <- (runif(N, 1, ncol(rj))) 
y <- (runif(N, 1, nrow(rj))) 

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Extract the colour at each pointt
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ind  <- round(y) * ncol(rj) + round(x)
cols <- rj[ind]
cols <- nara::packed_cols_to_hex_cols(cols)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Calculate the Voronoi
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
vor <- voronoi(as.double(x), as.double(y))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Plot all Voronoi polygons in their appropriate colour
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nr <- nara::nr_new(ncol(rj), nrow(rj))
for (i in seq_len(vor$polygons$npolygons)) {
  nr_polygon(nr, vor$polygons$individual[[i]]$x, vor$polygons$individual[[i]]$y, fill = cols[i])
}
grid.raster(nr)
```


## Voronoi Terminology

* Initial **sites**  (i.e. seed points)
* Fortune's algorithm returns
    * **vertices**
    * **segments**  (finite, semi-finite and infinite)
* This package reconstructs **polygons** (i.e. *Voronoi cells*) by 
  assembling the *segments* into polygons, and bounding any exterior
  *cells* by intersecting with a rectangular boundary


## Voronoi Tessellation feature comparison

| Package                   | rvoronoi | deldir | RTriangle |
|---------------------------|----------|--------|-----------|
| vertices                  |  Yes     | Yes    | Yes       |
| segments                  |  Yes     | Yes    | Yes       |
| interior polygons         |  Yes     | Yes    |           |
| bounded exterior polygons |  Yes     | Yes    |           |
| matched sites/polygons    |  Yes     | ???    |           |



## Delaunay Triangulation feature comparison

| Package                          | rvoronoi | deldir | RTriangle |
|----------------------------------|----------|--------|-----------|
| site indices (defining polygons) |  Yes     | Yes    |  Yes      |
| polygon coordinates              |  Yes     |        |           |



## Voronoi Tessellation

The following code calculates the Voronoi tessellation on 20 random points.


```{r voronoi}
library(rvoronoi)

set.seed(4)
x <- runif(20)
y <- runif(20)

vor <- voronoi(x, y) 

plot(vor, polys = FALSE, sites = TRUE, labels = TRUE, label_cex = 0.7)
plot(vor, labels = TRUE, label_cex = 0.7)
```


## Delaunay Triangulation

```{r delaunay}
library(rvoronoi)

set.seed(2024)
x <- runif(10)
y <- runif(10)

del <- delaunay(x, y)$segment

# Plot the sites
plot(x, y, asp = 1, col = 'red', ann = F, axes = F, pch = 19)

# Plot all finite segments.  
# This will not plot any of the segments which do not converge
segments(x[del$v1], y[del$v1], x[del$v2], y[del$v2])
segments(x[del$v3], y[del$v3], x[del$v2], y[del$v2])
segments(x[del$v1], y[del$v1], x[del$v3], y[del$v3])
```


## Delaunay Benchmark

Compare Delaunay Triangulation using

* `{rvoronoi}`
* `{delone}`  
    - `remotes::install_github("hypertidy/delone")`
* `{RTriangle}`
* `{deldir}`

<details>
<summary>Delaunay Triangulation Benchmark</summary>

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Delaunay
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
library(rvoronoi)
# library(delone)
library(RTriangle)
library(deldir)
#> deldir 2.0-4      Nickname: "Idol Comparison"
#> 
#>      The syntax of deldir() has changed since version 
#>      0.0-10.  Read the help!!!.

set.seed(1)
N <- 1000
x <- runif(N)
y <- runif(N)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# {rvoronoi}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
del_rvoronoi <- rvoronoi::delaunay(x, y)$segment

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# {delone} - normalise to 
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# del_delone <- delone::xy_tri(x, y)
# del_delone <- matrix(del_delone, ncol = 3, byrow = TRUE) |> as.data.frame()
# 
# stopifnot(identical(
#   rvoronoi:::normalise_del(del_rvoronoi),
#   rvoronoi:::normalise_del(del_delone)
# ))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# RTriangle
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
del_rtriangle <- RTriangle::triangulate(RTriangle::pslg(cbind(x, y)))$T
del_rtriangle <- as.data.frame(del_rtriangle)

stopifnot(identical(
  rvoronoi:::normalise_del(del_rvoronoi),
  rvoronoi:::normalise_del(del_rtriangle)
))


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# deldir
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
del_deldir <- deldir::deldir(x, y)
del_deldir <- deldir::triMat(del_deldir)
del_deldir <- as.data.frame(del_deldir)

stopifnot(identical(
  rvoronoi:::normalise_del(del_rvoronoi),
  rvoronoi:::normalise_del(del_deldir)
))
```

</details>

```{r}
bench::mark(
  # delone    = xy_tri  (x, y),
  rvoronoi  = delaunay(x, y, calc_polygons = TRUE),
  rvoronoi  = delaunay(x, y, calc_polygons = FALSE),
  rtriangle = triangulate(pslg(cbind(x, y))),
  deldir    = deldir(x, y),
  check = FALSE
)[,1:5]  |> knitr::kable()
```



# Voronoi Tessellation Benchmark

Compare:

* `{rvoronoi}`
* `{deldir}`
* `{RTriangle}`

<details>
<summary>Voronoi Tessellation Benchmark</summary>

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Voronoi
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
library(rvoronoi)
library(deldir)
library(RTriangle)

set.seed(1)
N <- 1000
x <- runif(N)
y <- runif(N)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# rvoronoi
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
vor_rvoronoi <- rvoronoi::voronoi(x, y)

if (FALSE) {
  plot(x, y, asp = 1, ann = F, axes = F, pch = '.')
  for (p in vor_rvoronoi$polygons) {
    polygon(p)
  }
}


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# deldir
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
vor_deldir   <- deldir::cvt(deldir::deldir(x, y), stopcrit = 'maxit', maxit = 1, verb = TRUE)

if (FALSE) {
  plot(x, y, asp = 1, ann = F, axes = F, pch = '.')
  for (p in vor_deldir$tiles) {
    polygon(p)
  }
}

if (FALSE) {
  plot(vor_deldir$tiles)
}
  
if (FALSE) {
  set.seed(42)
  x <- runif(20)
  y <- runif(20)
  dxy <- deldir(x,y,rw=c(0,1,0,1))
  cxy1 <- cvt(dxy,verb=TRUE, stopcrit = 'maxit', maxit = 1)
  plot(cxy1$tiles)
  with(cxy1$centroids,points(x,y,pch=20,col="red"))
  cxy2 <- cvt(dxy,stopcrit="m",verb=TRUE)
  plot(cxy2$tiles)
  with(cxy2$centroids,points(x,y,pch=20,col="red"))
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# RTriangle
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
tri <- triangulate(pslg(P = cbind(x, y)))
ve <- tri$VE |> as.data.frame()
ve <- subset(ve, V1 > 0 & V2 > 0)

if (FALSE) {
  plot(x, y, asp = 1, ann = F, axes = F, pch = '.')
  p1 <- tri$VP[ve$V1, ] |> as.data.frame()
  p2 <- tri$VP[ve$V2, ] |> as.data.frame()
  segments(p1$V1, p1$V2, p2$V1, p2$V2)
}
```

</details>



### Voronoi - vertices and segments only 

```{r}
bench::mark(
  `rvoronoi (match RTriangle)`  = voronoi(x, y, calc_polygons = FALSE),
  RTriangle = RTriangle::triangulate(pslg(P = cbind(x, y))),
  check = FALSE
)[,1:5] |> knitr::kable()
```


### Voronoi - polygons

```{r}
bench::mark(
  `rvoronoi (matched polygons)` = voronoi(x, y, calc_polygons = TRUE, match_sites = TRUE),
  `rvoronoi (unmatched polygons)` = voronoi(x, y, calc_polygons = TRUE, match_sites = FALSE),
  deldir    = deldir::cvt(deldir(x, y), stopcrit = 'maxit', maxit = 1),
  check = FALSE
)[,1:5] |> knitr::kable()
```




# Pathological Test Cases


## Pathological 1

* 100 points in a circle

```{r path1}
theta <- seq(0, 2*pi, length.out = 101)[-1]
x <- cos(theta)
y <- sin(theta)

vor <- voronoi(x, y)
plot(vor)
```




## Pathological 2

* 100 points in a circle
* 1 point at the centre

```{r path2}
theta <- seq(0, 2*pi, length.out = 101)[-1]
x <- c(0, cos(theta))
y <- c(0, sin(theta))

vor <- voronoi(x, y)
plot(vor)
```

## Pathological 2a

* 2 concentric circles (100 points each)
* 1 point at the centre

```{r path2a}
theta <- seq(0, 2*pi, length.out = 101)[-1]
x <- c(0, cos(theta), 2 * cos(theta))
y <- c(0, sin(theta), 2 * sin(theta))

vor <- voronoi(x, y)
plot(vor)
```


## Pathological 3

* 100 points in a line

```{r path3}
x <- seq(0, 2*pi, length.out = 100)
y <- 0.5 * x  

vor <- voronoi(x, y)
plot(vor)
```



## Pathological 4

* 2 concentric equilateral triangles 
* 1 point at the centre

```{r path4}
theta <- seq(0, 2*pi, length.out = 4)[-1]
x <- c(0, cos(theta), 2 * cos(theta))
y <- c(0, sin(theta), 2 * sin(theta))

vor <- voronoi(x, y)
plot(vor, labels = TRUE)
```


# Algorithms

* Fortune's Sweep Algorithm for Delaunay
    * Adapted original source code to call from R
* Polygon reconstruction
    * Use Voronoi vertices and segment connectivity to reconstruct polygonal Voronoi cells
    * Using "An optimal algorithm for extracting the regions of a plane graph" by
      Jiang & Bunke, Pattern Recognition Letters 14 (1993) pp553-558.
    * Adapted algorithm from binary search to instead use pre-indexed search bounds.
* Point in polygon (to match sites to Voronoi polygons)
    * Search for bounding box collision to filter polygon candidates for 
      exhaustive testing
    * Optimised point-in-polygon search with early termination possible (as Voronoi 
      polygons are **always** convex)
      Test edges using the "leftOf()" operator, and exit as soon as any 
      vertices are right of a polygon edge.
    * No benefit in building an acceleration structure for point-in-bounding-box
      when there of the order of ~100-1000 bounding boxes.  A linear search 
      (with flagging of polygons already claimed) is fast enough.
      If use case if Voronoi of >> 1000 points, then time taken to build acceleration structure
      (e.g. hierachical bounding boxes) would become worth it.

## Related 

* [mdsumner's helpful gist](https://gist.github.com/mdsumner/8db5ac01e47fa86f10e7ebc372e0ebda?permalink_comment_id=5171205)
* [`RTriangle::triangulate()`]()
* [`terra::voronoi()`](https://rspatial.github.io/terra/reference/voronoi.html)
* [`geos::geos_voronoi_polygons()`]()

